---
title: "Topic 7: Word Embeddings"
author: "Alex Vand"
date: "5/31/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(here)
library(tidytext)
library(tidyverse)
library(widyr)
library(irlba) #singular value decomposition
library(broom) # creating search_synonym function
library(textdata)
library(ggplot2)
library(dplyr)
```

```{r, include=FALSE}
data <- read_table(file = here("glove.6B.300d.txt"),
                   col_names = FALSE)
```


## 1.a Recreate the analyses in the last three chunks (find-synonyms, plot-synonyms, word-math) with the GloVe embeddings.

```{r, include=FALSE}
data <- data %>% 
  column_to_rownames(var = "X1")
  
```

```{r}
word_vectors <- as.matrix(data)
```

```{r}
search_synonyms <- function(word_vectors, selected_vector) {
dat <- word_vectors %*% selected_vector
    
similarities <- dat %>%
        tibble(token = rownames(dat), similarity = dat[,1])
similarities %>%
       arrange(-similarity) %>%
        select(c(2,3))
}
```


```{r}
fall <- search_synonyms(word_vectors,word_vectors["fall",])
slip <- search_synonyms(word_vectors,word_vectors["slip",])
```

```{r plot-synonyms}
slip %>%
    mutate(selected = "slip") %>%
    bind_rows(fall %>%
                  mutate(selected = "fall")) %>%
    group_by(selected) %>%
    top_n(15, similarity) %>%
    ungroup %>%
    mutate(token = reorder(token, similarity)) %>%
    ggplot(aes(token, similarity, fill = selected)) +
    geom_col(show.legend = FALSE) +
    facet_wrap(~selected, scales = "free") +
    coord_flip() +
    theme(strip.text=element_text(hjust=0, size=12)) +
    scale_y_continuous(expand = c(0,0)) +
    labs(x = NULL, title = "What word vectors are most similar to slip or fall?")
         
```


## 1.b How are they different from the embeddings created from the climbing accident data? Why do you think they are different?

These are different from the embeddings created from the climbing data given the particular meaning and context of "slip" and "fall" used here. "Fall" has a financial connotation while "slip" seems to be related to sports. Further comparison should include stemming the words of interest (removing fallen, falls, etc.).



## 2. Run the classic word math equation, "king" - "man" = ?

```{r}
king_man <- word_vectors["king",] - word_vectors["man",] 
search_synonyms(word_vectors, king_man)
```


## 3. Think of three new word math equations. They can involve any words you'd like, whatever catches your interest.



```{r}
bionic_vision <- word_vectors["bionic",] - word_vectors["vision",] 
search_synonyms(word_vectors, bionic_vision)
```

```{r}
beach_volleyball <- word_vectors["beach",] - word_vectors["volleyball",] 
search_synonyms(word_vectors, beach_volleyball)
```

```{r}
scuba_dive <- word_vectors["restaurant",] - word_vectors["menu",] 
search_synonyms(word_vectors, scuba_dive)
```


